---
title: 'Review of \texttt{R}'
author: "STA 360: Lab 1, Fall 2020 (This will not be graded or turned in)"
output: pdf_document
---

Today's agenda: A review of \texttt{R}, getting used to R markdown, vectors, matrices, scatterplots, and functions. 

***Lab Tasks***


1. Store three vectors using \textsf{rnorm()} of length $n = 100$ as \textsf{Var1}, \textsf{Var2}, and \textsf{Var3}.
```{r}
n <- 100
set.seed(9)
#Store 3 vectors using rnorm func
Var1 <- rnorm(n=100)
set.seed(341) #ensure you get a different draw each time
Var2 <- rnorm(n=100)
set.seed(142)
Var3 <- rnorm(n=100)

length(Var1)
head(Var1)
tail(Var1)
```


2. List all the items currently in the environment.

```{r}
ls()
```


3. Store \textsf{Var1} in a $10 \times 10$ matrix. Call this \textsf{myMatrix}.

```{r}

myMatrix <- matrix(data = Var1, nrow=10, ncol=10)
dim(myMatrix)
myMatrix

```


4. Create a scatterplot of \textsf{Var1} vs. \textsf{Var2}. On the same plotting window include histograms of \textsf{Var1} and \textsf{Var2}.

```{r}
par(mfrow=c(1,3))
plot(Var1, Var2, pch=16, col=rainbow(n))
hist(Var1)
hist(Var2)
```



5. Write a function that takes as its inputs, $p = 2$, $n$-dimensional vectors and a vector of length $p$ containing the names of these vectors. Your function combine these two vectors into a \textsf{data.frame()}, get the row-wise maximum and store this in a new vector. Finally produce a box-plot of this vector, store it as a separate .pdf, and return the mean value of this vector.

```{r}
mergeVectors <- function(v1, v2, myNames){
  myData <- data.frame(v1, v2)
  names(myData) <- myNames
  maxVector <- apply(myData, 1, max)
  pdf("myPlot.pdf")
  boxplot(maxVector)
  dev.off()
  final <- mean(maxVector)
  return(final)
  
  mergeVectors(Var1, Var2, c("myVar1", "myVar2"))
}
```


