---
title: "cameron-christopher-homework4.Rmd"
author: "Chris Cameron"
date: "9/15/2021"
header-includes:
   - \usepackage{hyperref, color}
   - \input{custom2}
output: pdf_document
indent: true
documentclass: article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. (10 points, 5 points each) Hoff, 3.10 (Change of variables). 

$$p_\psi (\psi) = p_\theta(h(\psi))\bigg|\frac{dh}{d\psi}\bigg| $$

a) $\theta \sim  beta(a,b)$ & $\psi = log[\theta/(1-\theta)]$. Obtain the form
of $p_\psi$ and plot it for the case that $a=b=1$ where $\theta = h(\psi)$.

$$\psi = log\bigg[\frac{\theta}{1-\theta}\bigg]$$
$$e^{\psi} = \frac{\theta}{1-\theta}$$
$$ \theta = e^{\psi}(1-\theta)$$
$$ \theta = e^{\psi}-\theta e^{\psi}$$
$$ e^{\psi}= \theta +\theta e^{\psi}$$
$$ e^{\psi}= \theta(1 +e^{\psi})$$
$$ \theta = \frac{e^{\psi}}{1 +e^{\psi}}$$
----------------------
$$\theta=:h(\psi)=\frac{e^\psi}{1+e^\psi}$$

$$\bigg|\frac{dh}{d\psi}\bigg| = \frac{e^{\psi}}{(1+e^\psi)^2} $$
----------------------------

$$ p_\theta (h(\psi)) = p_\theta (\theta) = \frac{1}{B(a,b)}[h(\psi)]^{a-1}[1-h(\psi)]^{b-1}$$
-----------------------------------------------------------------

$$ p_\psi (\psi) = p_\theta(h(\psi)) \bigg|\frac{dh}{d\psi}\bigg| = \frac{1}{B(a,b)}[h(\psi)]^{a-1}[1-h(\psi)]^{b-1}\frac{e^{\psi}}{(1 +e^{\psi})^2}$$
$$= \frac{1}{B(a,b)}\bigg[\frac{e^\psi}{1+e^\psi}\bigg]^{a-1}\bigg[\frac{1}{1+e^\psi}\bigg]^{b-1}\frac{e^{\psi}}{(1 +e^{\psi})^2} $$

$$= \frac{1}{B(a,b)}\bigg[\frac{e^\psi}{1+e^\psi}\bigg]^{a}\bigg[\frac{1}{1+e^\psi}\bigg]^{b}\frac{e^{\psi}}{(1 +e^{\psi})^2}\bigg(\frac{e^\psi}{1+e^\psi}\bigg)\bigg(\frac{1}{1+e^\psi}\bigg) $$

$$ p_{\psi} (\psi) = \frac{1}{B(a,b)}\bigg[\frac{e^\psi}{1+e^\psi}\bigg]^{a}\bigg[\frac{1}{1+e^\psi}\bigg]^{b}$$

```{r}
set.seed(1)
theta.sim <- seq(from = -0.5, to = 1.5, length.out = 1000)
a <- 1
b <- 1

y <- (gamma(a+b)/gamma(a)*gamma(b))*(((exp(log((theta.sim)/(1-theta.sim))))/(1+exp(log((theta.sim)/(1-theta.sim)))))^a)*(((1)/(1+exp(log((theta.sim)/(1-theta.sim)))))^b)

#y[is.nan(y)] <- 0

plot(theta.sim, y, type = "l", main = "Probability Distribution", 
     xlab = "psi", yaxt = "n", ylab = "likelihood") 
```

b) Let $\theta \sim gamma(a,b)$ and let $\psi = log(\theta)$. Obtain the form of $p_\psi$ and plot it for the case that $a=b=1$.

$$p_\psi (\psi) = p_\theta(h(\psi))\bigg|\frac{dh}{d\psi}\bigg| $$

-----------------------------------

$$log(\theta) = \psi $$
$$\theta = e^\psi $$
-----------------------------------

$$\theta=:h(\psi)=e^\psi$$

$$\bigg|\frac{dh}{d\psi}\bigg| = e^\psi $$
-----------------------------------

$$p_\theta (h(\psi)) = p_\theta (\theta) = \frac{b^a}{\Gamma(a)}\theta^{a-1}e^{-b\theta}$$
$$p_\theta (h(\psi)) = p_\theta (\theta) = \frac{b^a}{\Gamma(a)}(e^\psi)^{a-1}e^{-b(e^\psi)}$$
$$p_\theta (h(\psi)) = p_\theta (\theta) = \frac{b^a}{\Gamma(a)}e^{\psi a-\psi}e^{-be^\psi}$$
$$p_\theta (h(\psi)) = p_\theta (\theta) = \frac{b^a}{\Gamma(a)}e^{\psi a-\psi-be^\psi}$$

-----------------------------------

$$p_\psi (\psi) = p_\theta(h(\psi)) \bigg|\frac{dh}{d\psi}\bigg| = \frac{b^a}{\Gamma(a)}e^{\psi a-\psi-be^\psi}e^\psi $$
$$p_\psi (\psi) = \frac{b^a}{\Gamma(a)}e^{\psi a-be^\psi}$$

```{r}
theta.sim <- seq(from = -10, to = 5, length.out = 1000)
a <- 1
b <- 1

y <- ((b^a)/(gamma(a)))*(exp((theta.sim*a)-(b*exp(theta.sim))))

y[is.nan(y)] <- 0

plot(theta.sim, y, type = "l", main = "Probability Distribution", 
     xlab = "psi", yaxt = "n", ylab = "likelihood") 
```

(25 points total) Please refer to lab 4 and complete tasks 4---5. 

```{r}
prior = data.frame(m = 0, c = 1, a = 0.5, b = 50)
```


# Task 4

```{r}
# Number of posterior simulations
sim = 100000

# initialize vectors to store samples
mus = NULL
lambdas = NULL
muc = NULL
lambdac = NULL

# Following formula from the NormalGamma with 
# the update paramaters accounted accounted for below 

lambdas = rgamma(sim, shape = postS$a, rate = postS$b)
lambdac = rgamma(sim, shape = postC$a, rate = postC$b)


mus = sapply(sqrt(1/(postS$c*lambdas)),rnorm, n = 1, mean = postS$m)
muc = sapply(sqrt(1/(postC$c*lambdac)),rnorm, n = 1, mean = postC$m)

# Store simulations
simDF = data.frame(lambda = c(lambdas, lambdac),
                   mu = c(mus, muc),
                   Treatment = rep(c("Spurters", "Controls"),
                                   each = sim))

simDF$lambda = simDF$lambda^{-0.5}



sgtc = 0 #Stands for "spurters greater than control"

#Calculate proportion of spurters with greater change in IQ than the control group
for(i in 1:100000){
  if(simDF[i, "mu"] > simDF[i+100000, "mu"]){
    sgtc <- sgtc+1
  }
}

print(sgtc/100000)
```

There is a predicted 97.111% chance that a arndomly selected spurter will experience
a greater increase in their IQ score over a year than a randomly selected child
from the control group.

# Task 5

```{r}

# Following formula from the NormalGamma with 
# the prior parameters accounted accounted for below 

lambda_prior = rgamma(1000, shape = prior$a, rate = prior$b)

mu_prior = sapply(sqrt(1/(prior$c*lambda_prior)),rnorm, n = 1, mean = prior$m)

# Store simulations
simDF = data.frame(lambda = lambda_prior,
                   mu = mu_prior)

simDF$lambda = simDF$lambda^{-0.5}

# Plot the simulations
ggplot(data = simDF, aes(x = mu, y = lambda)) +
  geom_point(alpha = 0.2) + 
  labs(x = expression(paste(mu, " (Mean Change in IQ Score)")),
       y = expression(paste(lambda^{-1/2}, " (Std. Dev. of Change)")))  + 
  ggtitle("Posterior Samples")+ 
  theme(plot.title = element_text(hjust = 0.5))
```

