---
title: "Homework 7"
author: "Chris Cameron"
date: "11/10/2021"
header-includes:
   - \usepackage{hyperref, color}
   - \input{custom2}
output: pdf_document
indent: true
documentclass: article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# A)

```{r}
bluecrab <- read.csv("data/bluecrab.dat", sep = ' ', header = FALSE)
orangecrab <- read.csv("data/orangecrab.dat", sep = ' ', header = FALSE)
```


```{r}
#Set Prior Parameters
mu_blue <- apply(bluecrab,2,mean)
mu_orange <- apply(orangecrab,2,mean)
L_blue <- cov(bluecrab)
L_orange <- cov(orangecrab)
S_blue = L_blue
S_orange = L_orange
nu_blue <- 7
nu_orange <- 7
iterations <- 10000
```

```{r}
library (mvtnorm)
library(MCMCpack)
```

```{r}
#rbind does not seem to be (properly) returnable in a list,
#which is why I am not using a function

mu <- mu_blue
L <- L_blue
S <- S_blue
nu <- nu_blue
Y <- bluecrab

n <- dim (Y)[1]
Sigma <- S #initialize to prior?

THETA_blue <- SIGMA_blue <- NULL
set.seed(6)

for(s in 1:iterations){
  
  #update theta
  Ln <- solve(solve(L) + n*solve(Sigma))
  mun <- Ln %*% (solve(L) %*% mu + n*solve(Sigma) %*% mu) #mu is equal to mean
  theta <- rmvnorm(1, mu, L)
  
  #update sigma
  Sn <- S + (t(Y) - c(theta)) %*% t(t(Y)-c(theta))
  
  Sigma <- solve(rwish(nu + n, solve(Sn)))

  #Save Results
  THETA_blue <- rbind(THETA_blue, theta)
  SIGMA_blue <- rbind(SIGMA_blue, c(Sigma))
}
```

```{r}
mu <- mu_orange
L <- L_orange
S <- S_orange
nu <- nu_orange
Y <- orangecrab

n <- dim (Y)[1]
Sigma <- S #initialize to prior?

THETA_orange <- SIGMA_orange <- NULL

for(s in 1:iterations){
  
  #update theta
  Ln <- solve(solve(L) + n*solve(Sigma))
  mun <- Ln %*% (solve(L) %*% mu + n*solve(Sigma) %*% mu) #mu is equal to mean
  theta <- rmvnorm(1, mu, L)
  
  #update sigma
  Sn <- S + (t(Y) - c(theta)) %*% t(t(Y)-c(theta))
  
  Sigma <- solve(rwish(nu + n, solve(Sn)))

  #Save Results
  THETA_orange <- rbind(THETA_orange, theta)
  SIGMA_orange <- rbind(SIGMA_orange, c(Sigma))
}
```

# B)
```{r}
plot(THETA_blue, col = rgb(red = 0, green = 0.5, blue = 1, alpha = 0.1),
     main = "Dimensions of Crabs", xlab = "Body Depth", ylab = "Rear Width",
     xlim = c(0, 22), ylim = c(0, 29))
points(THETA_orange, col = rgb(red = 1, green = 0.5, blue = 0, alpha = 0.1))
legend("topleft", c("Blue Crab", "Orange Crab"), fill = c("deepskyblue", "orange"))
#Citation: https://stackoverflow.com/questions/14683766/
#adding-key-legend-to-multi-histogram-plot-in-r
```

```{r, echo = FALSE}
print("Average Blue Crab Body Depth")
mean(THETA_blue[1])
print("Average Blue Crab Rear Width")
mean(THETA_blue[2])
print("Average Orange Crab Body Depth")
mean(THETA_orange[1])
print("Average Orange Crab Rear Width")
mean(THETA_orange[2])

```

The above plot demonstrates that both orange and blue crabs demonstrate a positive
linear relationship between their rear widths and body depths. The relationship
between width and depth appears to have a similar slope for both species, 
with an orange crab having a larger rear width than a blue crab of equal body depth.
Furthermore, orange crabs have rear widths that are 2.1 cm larger than that of
blue crabs, and body depths that are 0.2 cm larger, demonstrating that orange
crabs are larger on average.

# C)

```{r}
blue_cor <- list()
for (x in 1:iterations){
  blue_cor[x] <- cov2cor(matrix(SIGMA_blue[x,1:4], nrow=2, ncol=2))[2]
}

orange_cor <- list()
for (x in 1:iterations){
  orange_cor[x] <- cov2cor(matrix(SIGMA_orange[x,1:4], nrow=2, ncol=2))[2]
}
```

```{r}
plot(density(unlist(blue_cor)),
     main = "Density of Correlation Between Width & Depth For Blue Crabs")
abline(v = mean(unlist(blue_cor)))

plot(density(unlist(orange_cor)),
     main = "Density of Correlation Between Width & Depth For Orange Crabs")
abline(v = mean(unlist(orange_cor)))


```

```{r}
a = 0
for(i in 1:iterations){
  if(unlist(orange_cor)[i] > unlist(blue_cor)[i]){
    a <- a+1
  }
}
print(a/iterations)

```

```{r}
t.test(unlist(orange_cor), unlist(blue_cor), alternative = "less")
```

Both types of crabs have correlations between their body depths and rear widths
that are very high, with the density of this correlation being skewed left for
both. The average correlation for orange crabs is .985, while that for blue 
crabs is .964. The average correlation for any one sample of orange crabs has
a probability of .789 of being higher than that of any one sample of blue crabs.
The probability that the overall mean correlation of orange crabs is higher than
that of blue crabs, using a t-test with 13460 degrees of freedom, is nearly 1.
These values suggest that the correlation between correlation between body depth
and rear width is on average higher in orange crabs than in blue crabs.
